Appendix 2: Simulating 20-species landscapes

### 2016-02-18 Harris_2015/show-results.Rmd
many changes

# Sort a vector of alphanumeric strings by the numeric component
# as if they were integers.  For example, V20 is larger than V12,
# even though V12 comes first alphabetically
alnum_sort = function(x){
  raw = as.integer(gsub("[[:alpha:]]", "", x))
  x[order(raw)]
}

@@

    # Pull out the corresponding chunk of the "x" data frame, based on n_sites and rep_name
    x_subset = x[x$n_sites == n_sites & x$rep_name == rep_name & x$method == "correlation", ]

@@

## Calculate model performance:

@@

result_summary$method = reorder(result_summary$method, -result_summary$r2)

result_summary$simulation_type = reorder(result_summary$simulation_type, -result_summary$r2)

result_summary = result_summary %>% 
  group_by(method) %>% 
  summarise(mean_r2 = round(100 * mean(r2))) %>%
  mutate(method_r2 = paste0(method, " (0.", mean_r2, ")")) %>%
  select(method, method_r2) %>%
  inner_join(result_summary, "method")

result_summary$simulation_type_long = plyr::revalue(
  result_summary$simulation_type,
  c(no_env = "constant environment",
    env = "heterogeneous environment",
    abund = "abundance")
)

@@

#### Model performance with no environmental variation:

```{r}
result_summary %>% 
  filter(simulation_type == "no_env") %>%
  group_by(method) %>% 
  summarise(mean(r2)) %>% 
  kable(digits = 3)
```

#### Model performance with environmental heterogeneity:

```{r}
result_summary %>% 
  filter(simulation_type == "env") %>%
  group_by(method) %>% 
  summarise(mean(r2)) %>% 
  kable(digits = 3)
```

#### Model performance with per-capita species interactions:

```{r}
result_summary %>% 
  filter(simulation_type == "abund") %>%
  group_by(method) %>% 
  summarise(mean(r2)) %>% 
  kable(digits = 3)
```

## Plot the R-squared results (Figure 3)

@@

pdf("manuscript-materials/figures/performance.pdf", width = 8, height = 2.5)
ggplot(result_summary, aes(x = n_sites, y = r2, col = method_r2, shape = method_r2)) + 
  facet_grid(~simulation_type_long) + 
  geom_line(size = .5) + 
  geom_point(size = 2.5, fill = "white") + 
  scale_shape_manual(values = c(16, 22, 17, 23, 18, 24, 15), name = legend_name) +
  geom_hline(yintercept = 0, size = 1/2) +
  geom_vline(xintercept = 0, size = 1) + 
  #scale_shape_manual(values = c(21, 25, 15, 18)) + 
  coord_cartesian(ylim = c(-.01, 0.76)) + 
  ylab(expression(R^2)) + 
  xlab("Number of sites (log scale)") + 
  scale_x_log10(breaks = unique(x$n_sites), limits = range(x$n_sites)) +
  theme_bw(base_size = 11) + 
  theme(panel.margin = grid::unit(1.25, "lines")) + 
  theme(panel.border = element_blank(), axis.line = element_blank()) + 
  theme(
    panel.grid.minor = element_blank(), 
    panel.grid.major.y = element_line(color = "lightgray", size = 1/4), 
    panel.grid.major.x = element_blank()
  ) + 
  theme(strip.background = element_blank(), legend.key = element_blank()) + 
  theme(plot.margin = grid::unit(c(.01, .01, .75, .1), "lines")) + 
  theme(
    axis.title.x = element_text(vjust = -0.2, size = 12), 
    axis.title.y = element_text(angle = 0, hjust = -.1, size = 12)
  ) + 
  scale_color_brewer(palette = "Dark2", name = legend_name)
dev.off()
```

@@

summary(
  lmer(
    r2 ~ method + n_sites + simulation_type + (1|rep_name), 
    data = landscape_estimates
  )
)
```

## Summarize the inferential results of the Markov network and the null model:

@@

# Pairs's Z score is based on C-scores, which are positive when species are 
# disaggregated.  So significantly positive interactions == negative estimates
pairs_summary$sig_pos = pairs_summary$estimate < qnorm(.025)
pairs_summary$sig_neg = pairs_summary$estimate > qnorm(.975)

# Markov network is significant when lower bound is above zero or lower bound is
# below zero
markov_summary$sig_pos = markov_summary$lower > 0
markov_summary$sig_neg = markov_summary$upper < 0
```

@@

# Function for smoothing the error rates with a generalized additive model
# and plotting the results
my_geom_smooth = function(data, color, ...){
  geom_smooth(
    data = data,
    aes(
      x = truth, 
      y = as.integer((sig_neg & truth > 0) | (sig_pos & truth < 0)),
      color = color
    ),
    method = gam, 
    family = binomial, 
    formula = y ~ s(x),
    se = FALSE,
    n = 1024,
    ...
  )
}

@@

error_smoother = function(data){
  predict(
    gam(
      I((sig_neg & truth > 0) | (sig_pos & truth < 0)) ~ s(truth),
      data = data,
      family = binomial
    ),
    data.frame(truth = truth_seq),
    type = "response"
  )
}

y_pairs = error_smoother(pairs_summary)
y_markov = error_smoother(markov_summary)

@@

```{r, cache = TRUE}
# Coverage versus true beta value.  The top and bottom 0.5%
# of the distribution have been omitted to prevent bad behavior
# by the smoother in the tails.
markov_summary %>%
  filter(percent_rank(truth) > .005 & percent_rank(truth) < .995) %>%
  mutate(covered = truth > lower & truth < upper) %>%
  ggplot(aes(x = truth, y = as.integer(covered))) + 
  facet_grid(~simulation_type) + 
  geom_smooth(method = gam, formula = y ~ s(x), family = binomial) + 
  theme_bw() + 
  coord_cartesian(ylim = c(0, 1)) + 
  geom_vline(xintercept = 0) + 
  geom_hline(yintercept = 0.95, color = "red") + 
  ylab("Coverage")

>>>>

# Sort a vector of alphanumeric strings by the numeric component
# as if they were integers.  For example, V20 should be larger 
# than V12, even though V12 comes first alphabetically
alnum_sort = function(x){
  raw = as.integer(gsub("[[:alpha:]]", "", x))
  x[order(raw)]
}

@@

    # Pull out the corresponding chunk of the "x" data frame, based on n_sites 
    # and rep_name
    is_correct_sim = x$n_sites == n_sites & x$rep_name == rep_name
    x_subset = x[is_correct_sim & x$method == "correlation", ]

@@

## Calculate model performance (R-squared):

For each combination of `method` and `simulation_type`, fit a linear model.
Then, for each combination of `method`, `simulation_type`, and `n_sites`,
report the proportion of variance explained by the corresponding linear model.

@@

# Set the ordering of the data frame based on R-squared
result_summary$method = reorder(result_summary$method, -result_summary$r2)
result_summary$simulation_type = reorder(
  result_summary$simulation_type, 
  -result_summary$r2
)

# Add a column that includes method name and its mean R-squared,
# for plotting purposes below
result_summary = result_summary %>% 
  group_by(method) %>% 
  summarise(mean_r2 = round(100 * mean(r2))) %>%
  mutate(method_r2 = paste0(method, " (0.", mean_r2, ")")) %>%
  select(method, method_r2) %>%
  inner_join(result_summary, "method")

# Rename the simulation types for clearer graph labels
result_summary$simulation_type_long = plyr::revalue(
  result_summary$simulation_type,
  c(no_env = "constant environment",
    env = "heterogeneous environment",
    abund = "abundance")
)

@@

#### Calculate average R-squared across methods and simulation types

```{r}
result_summary %>% 
  group_by(method, simulation_type) %>% 
  summarise(mean(r2)) %>% 
  spread(simulation_type, `mean(r2)`) %>%
  kable(digits = 3)
```

## Save the finer-grained R-squared results as Figure 3

@@

pdf("manuscript-materials/figures/performance.pdf", width = 8, height = 2.5)
ggplot(result_summary, aes(x = n_sites, y = r2, col = method_r2, shape = method_r2)) + 
  facet_grid(~simulation_type_long) + 
  geom_line(size = .5) + 
  geom_point(size = 2.5, fill = "white") + 
  scale_shape_manual(values = c(16, 22, 17, 23, 18, 24, 15), name = legend_name) +
  geom_hline(yintercept = 0, size = 1/2) +
  geom_vline(xintercept = 0, size = 1) + 
  coord_cartesian(ylim = c(-.01, 0.76)) + 
  ylab(expression(R^2)) + 
  xlab("Number of sites (log scale)") + 
  scale_x_log10(breaks = unique(x$n_sites), limits = range(x$n_sites)) +
  theme_bw(base_size = 11) + 
  theme(panel.margin = grid::unit(1.25, "lines")) + 
  theme(panel.border = element_blank(), axis.line = element_blank()) + 
  theme(
    panel.grid.minor = element_blank(), 
    panel.grid.major.y = element_line(color = "lightgray", size = 1/4), 
    panel.grid.major.x = element_blank()
  ) + 
  theme(strip.background = element_blank(), legend.key = element_blank()) + 
  theme(plot.margin = grid::unit(c(.01, .01, .75, .1), "lines")) + 
  theme(
    axis.title.x = element_text(vjust = -0.2, size = 12), 
    axis.title.y = element_text(angle = 0, hjust = -.1, size = 12)
  ) + 
  scale_color_brewer(palette = "Dark2", name = legend_name)
dev.off()
```

@@

summary(
  lmer(
    r2 ~ method + n_sites + simulation_type + (1|rep_name), 
    data = landscape_estimates
  ),
  correlation = FALSE
)
```

## Identify statistical significance for the Markov network and Pairs

For Pairs, compare the Z-scores with Gaussian quantiles for 95% coverage;
for the Markov network, use the approximate confidence intervals estimated in
Appendix 3.

@@

# Pairs's Z score is based on C-scores, which are positive when species are 
# disaggregated.  So significantly negative scores imply positive interactions
# and vice versa
pairs_summary$sig_pos = pairs_summary$estimate < qnorm(.025)
pairs_summary$sig_neg = pairs_summary$estimate > qnorm(.975)

# Markov network is significant when lower bound is above zero or lower bound is
# below zero
markov_summary$sig_pos = markov_summary$lower > 0
markov_summary$sig_neg = markov_summary$upper < 0

confidently_right = function(data){
  (data$sig_pos & data$truth > 0) | (data$sig_neg & data$truth < 0)
}

confidently_wrong = function(data){
  (data$sig_pos & data$truth < 0) | (data$sig_neg & data$truth > 0)
}
```

@@

@@

error_smoother = function(data, f){
  predict(
    gam(
      f(data) ~ s(truth),
      data = data,
      family = binomial
    ),
    data.frame(truth = truth_seq),
    type = "response"
  )
}

p_right_pairs = error_smoother(pairs_summary, confidently_right)
p_wrong_pairs = error_smoother(pairs_summary, confidently_wrong)

p_right_markov = error_smoother(markov_summary, confidently_right)
p_wrong_markov = error_smoother(markov_summary, confidently_wrong)

par(mfrow = c(1, 2))
plot(truth_seq, p_right_pairs + p_wrong_pairs, type = "l", ylim = c(0, 1), yaxs = "i")
lines(truth_seq, p_right_markov + p_wrong_markov, col = 2)
abline(h = 0.05)

plot(truth_seq, p_wrong_pairs, type = "l", ylim = c(0, 1), yaxs = "i")
lines(truth_seq, p_wrong_markov, col = 2)
par(mfrow = c(1, 1))

y_markov = error_smoother(markov_summary, confidently_wrong)

@@

```{r, cache = TRUE}
# Coverage versus true beta value.  The top and bottom 0.5%
# of the distribution have been omitted to prevent bad behavior
# by the smoother in the tails.
markov_summary %>%
  filter(percent_rank(truth) > .005 & percent_rank(truth) < .995) %>%
  mutate(covered = truth > lower & truth < upper) %>%
  ggplot(aes(x = truth, y = as.integer(covered))) + 
  facet_grid(~simulation_type) + 
  geom_smooth(method = gam, formula = y ~ s(x), method.args = list(family = "binomial")) + 
  theme_bw() + 
  coord_cartesian(ylim = c(0, 1)) + 
  geom_vline(xintercept = 0) + 
  geom_hline(yintercept = 0.95, color = "red") + 
  ylab("Coverage")

### 2016-02-18 Harris_2015/show-results.tex
many changes

    \CommentTok{# Pull out the corresponding chunk of the "x" data frame, based on n_sites and rep_name}
    \NormalTok{x_subset =}\StringTok{ }\NormalTok{x[x$n_sites ==}\StringTok{ }\NormalTok{n_sites &}\StringTok{ }\NormalTok{x$rep_name ==}\StringTok{ }\NormalTok{rep_name &}\StringTok{ }\NormalTok{x$method ==}\StringTok{ "correlation"}\NormalTok{, ]}

@@

\NormalTok{result_summary$simulation_type =}\StringTok{ }\KeywordTok{reorder}\NormalTok{(result_summary$simulation_type, -result_summary$r2)}

@@

\paragraph{Model performance with no environmental
variation:}\label{model-performance-with-no-environmental-variation}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{result_summary %>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(simulation_type ==}\StringTok{ "no_env"}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(method) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\KeywordTok{mean}\NormalTok{(r2)) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{kable}\NormalTok{(}\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lr@{}}
\toprule
method & mean(r2)\tabularnewline
\midrule
\endhead
Markov network & 0.525\tabularnewline
GLM & 0.472\tabularnewline
partial correlation & 0.403\tabularnewline
partial BayesComm & 0.394\tabularnewline
correlation & 0.291\tabularnewline
null & 0.227\tabularnewline
BayesComm & 0.206\tabularnewline
\bottomrule
\end{longtable}

\paragraph{Model performance with environmental
heterogeneity:}\label{model-performance-with-environmental-heterogeneity}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{result_summary %>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(simulation_type ==}\StringTok{ "env"}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(method) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\KeywordTok{mean}\NormalTok{(r2)) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{kable}\NormalTok{(}\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lr@{}}
\toprule
method & mean(r2)\tabularnewline
\midrule
\endhead
Markov network & 0.451\tabularnewline
GLM & 0.405\tabularnewline
partial correlation & 0.322\tabularnewline
partial BayesComm & 0.302\tabularnewline
correlation & 0.183\tabularnewline
null & 0.125\tabularnewline
BayesComm & 0.110\tabularnewline
\bottomrule
\end{longtable}

\paragraph{Model performance with per-capita species
interactions:}\label{model-performance-with-per-capita-species-interactions}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{result_summary %>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(simulation_type ==}\StringTok{ "abund"}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{group_by}\NormalTok{(method) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\KeywordTok{mean}\NormalTok{(r2)) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{kable}\NormalTok{(}\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lr@{}}
\toprule
method & mean(r2)\tabularnewline
\midrule
\endhead
Markov network & 0.384\tabularnewline
GLM & 0.283\tabularnewline
partial correlation & 0.200\tabularnewline
partial BayesComm & 0.166\tabularnewline
correlation & 0.117\tabularnewline
null & 0.075\tabularnewline
BayesComm & 0.060\tabularnewline
\bottomrule
\end{longtable}

\subsection{Plot the R-squared results (Figure
3)}\label{plot-the-r-squared-results-figure-3}

@@

\subsection{Summarize the inferential results of the Markov network and
the null
model:}\label{summarize-the-inferential-results-of-the-markov-network-and-the-null-model}

@@

\begin{verbatim}
## geom_smooth: method="auto" and size of largest group is >=1000, so using gam with formula: y ~ s(x, bs = "cs"). Use 'method = x' to change the smoothing method.
\end{verbatim}

\includegraphics{show-results_files/figure-latex/unnamed-chunk-13-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Coverage versus true beta value.  The top and bottom 0.5%}
\CommentTok{# of the distribution have been omitted to prevent bad behavior}
\CommentTok{# by the smoother in the tails.}
\NormalTok{markov_summary %>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{percent_rank}\NormalTok{(truth) >}\StringTok{ }\NormalTok{.}\DecValTok{005} \NormalTok{&}\StringTok{ }\KeywordTok{percent_rank}\NormalTok{(truth) <}\StringTok{ }\NormalTok{.}\DecValTok{995}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{covered =} \NormalTok{truth >}\StringTok{ }\NormalTok{lower &}\StringTok{ }\NormalTok{truth <}\StringTok{ }\NormalTok{upper) %>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{truth, }\DataTypeTok{y =} \KeywordTok{as.integer}\NormalTok{(covered))) +}\StringTok{ }
\StringTok{  }\KeywordTok{facet_grid}\NormalTok{(~simulation_type) +}\StringTok{ }
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\DataTypeTok{method =} \NormalTok{gam, }\DataTypeTok{formula =} \NormalTok{y ~}\StringTok{ }\KeywordTok{s}\NormalTok{(x), }\DataTypeTok{family =} \NormalTok{binomial) +}\StringTok{ }
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() +}\StringTok{ }
\StringTok{  }\KeywordTok{coord_cartesian}\NormalTok{(}\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)) +}\StringTok{ }
\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =} \DecValTok{0}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{geom_hline}\NormalTok{(}\DataTypeTok{yintercept =} \FloatTok{0.95}\NormalTok{, }\DataTypeTok{color =} \StringTok{"red"}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{"Coverage"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{show-results_files/figure-latex/unnamed-chunk-14-1.pdf}

>>>>

    \CommentTok{# Pull out the corresponding chunk of the "x" data frame, based on n_sites }
    \CommentTok{# and rep_name}
    \NormalTok{is_correct_sim =}\StringTok{ }\NormalTok{x$n_sites ==}\StringTok{ }\NormalTok{n_sites &}\StringTok{ }\NormalTok{x$rep_name ==}\StringTok{ }\NormalTok{rep_name}
    \NormalTok{x_subset =}\StringTok{ }\NormalTok{x[is_correct_sim &}\StringTok{ }\NormalTok{x$method ==}\StringTok{ "correlation"}\NormalTok{, ]}

@@

\NormalTok{result_summary$simulation_type =}\StringTok{ }\KeywordTok{reorder}\NormalTok{(}
  \NormalTok{result_summary$simulation_type, }
  \NormalTok{-result_summary$r2}
\NormalTok{)}

@@

\paragraph{Calculate average R-squared across methods and simulation
types}\label{calculate-average-r-squared-across-methods-and-simulation-types}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{result_summary %>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(method, simulation_type) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{summarise}\NormalTok{(}\KeywordTok{mean}\NormalTok{(r2)) %>%}\StringTok{ }
\StringTok{  }\KeywordTok{spread}\NormalTok{(simulation_type, }\StringTok{`}\DataTypeTok{mean(r2)}\StringTok{`}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{kable}\NormalTok{(}\DataTypeTok{digits =} \DecValTok{3}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[c]{@{}lrrr@{}}
\toprule
method & no\_env & env & abund\tabularnewline
\midrule
\endhead
Markov network & 0.525 & 0.451 & 0.384\tabularnewline
GLM & 0.472 & 0.405 & 0.283\tabularnewline
partial correlation & 0.403 & 0.322 & 0.200\tabularnewline
partial BayesComm & 0.394 & 0.302 & 0.166\tabularnewline
correlation & 0.291 & 0.183 & 0.117\tabularnewline
null & 0.227 & 0.125 & 0.075\tabularnewline
BayesComm & 0.206 & 0.110 & 0.060\tabularnewline
\bottomrule
\end{longtable}

\subsection{Save the R-squared results as Figure
3}\label{save-the-r-squared-results-as-figure-3}

@@

\subsection{Identify statistical significance for the Markov network and
Pairs}\label{identify-statistical-significance-for-the-markov-network-and-pairs}

@@

\includegraphics{show-results_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Coverage versus true beta value.  The top and bottom 0.5%}
\CommentTok{# of the distribution have been omitted to prevent bad behavior}
\CommentTok{# by the smoother in the tails.}
\NormalTok{markov_summary %>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\KeywordTok{percent_rank}\NormalTok{(truth) >}\StringTok{ }\NormalTok{.}\DecValTok{005} \NormalTok{&}\StringTok{ }\KeywordTok{percent_rank}\NormalTok{(truth) <}\StringTok{ }\NormalTok{.}\DecValTok{995}\NormalTok{) %>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{covered =} \NormalTok{truth >}\StringTok{ }\NormalTok{lower &}\StringTok{ }\NormalTok{truth <}\StringTok{ }\NormalTok{upper) %>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =} \NormalTok{truth, }\DataTypeTok{y =} \KeywordTok{as.integer}\NormalTok{(covered))) +}\StringTok{ }
\StringTok{  }\KeywordTok{facet_grid}\NormalTok{(~simulation_type) +}\StringTok{ }
\StringTok{  }\KeywordTok{geom_smooth}\NormalTok{(}\DataTypeTok{method =} \NormalTok{gam, }\DataTypeTok{formula =} \NormalTok{y ~}\StringTok{ }\KeywordTok{s}\NormalTok{(x), }\DataTypeTok{method.args =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{family =} \StringTok{"binomial"}\NormalTok{)) +}\StringTok{ }
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() +}\StringTok{ }
\StringTok{  }\KeywordTok{coord_cartesian}\NormalTok{(}\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{1}\NormalTok{)) +}\StringTok{ }
\StringTok{  }\KeywordTok{geom_vline}\NormalTok{(}\DataTypeTok{xintercept =} \DecValTok{0}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{geom_hline}\NormalTok{(}\DataTypeTok{yintercept =} \FloatTok{0.95}\NormalTok{, }\DataTypeTok{color =} \StringTok{"red"}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{ylab}\NormalTok{(}\StringTok{"Coverage"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{show-results_files/figure-latex/unnamed-chunk-12-1.pdf}

### 2016-02-18 Harris_2015/simulation.Rmd
many changes

@@

I used the following R function to generate "true" parameters with the method above, 
then simulate a presence-absence landscape based on those parameters using Markov chain Monte Carlo, and finally save the results to a "fakedata" folder.

I used Gibbs sampling as my Markov chain Monte Carlo transition operator.
In each round of Gibbs sampling, I cycled through all the species, randomly
updating each one's presence/absence vector in response to its conditional
occurrence probability:

@@

where the logistic function is $\frac{1}{1+e^{-x}}$.

For the abundance-based simulations, abundance was simulated with a Poisson
distribution with rate parameter given by 
$\mathrm{log}(1 + exp(\alpha_i + \sum_j\beta_{ij}y_j))$.

@@

  # Landscape starts as if betas were all zero. Each species' occurrence probability
  # or abundance depends on its alpha value and on the environment (assuming alpha_env 
  # is not zero).
  x = matrix(
    f(rep(1, n_sites) %*% t(alpha) + env %*% alpha_env), 
    nrow = n_sites, 
    ncol = n_spp
  )

@@

Loop to simulate and save all the landscapes:

>>>>

The $alpha$ coefficients controlling species' intercepts were drawn from 
Gaussian distributions with unit variance and a mean given by `mean_alpha`.
In the R code at the end of this document, `mean_alpha` was defined as $-1$ for presence-absence simulations and $5$ for the abundance simulations.

The $\beta$ coefficients' magnitudes were drawn from an exponential 
distribution with rate parameter equal to 1.  The sign of each coefficient was
randomly sampled, based on the value of `p_neg`, defined below. as 0.75 (75%
negative) for the presence-absence simulations and 1.0 (100% negative) for the
simulations that included abundance.

@@

The following function calls the function above to generate "true" parameters, 
then uses them to simulate a landscape with Markov chain Monte Carlo, and 
finally saves the results to a "fakedata" folder.
I used Gibbs sampling as my Markov chain Monte Carlo transition operator.
In each round of Gibbs sampling, I cycled through all the species, randomly
updating each one's binary presence/absence vector with random samples from
a Bernoulli distribution whose conditional probability was defined by

@@

where the logistic function is $\frac{1}{1+e^{-x}}$. For efficiency, 
$\sum_j\beta_{ij}y_j$ is calculated with matrix multiplication in the code
below.

For simulations that included environmental heterogeneity, $\alpha_i$ varied 
from site to site as follows:

* `n_env` environmental variables were simulated for each site; these were
drawn from zero-mean Gaussians with standard deviation given by `sd`. In the 
R code at the end of this document, both `n_env` and `sd` were set to 2, 
corresponding to an environment with two important axes of variation and 
had a standard deviation of two.

* The local value of $\alpha_i$ for species $i$ is given by the species-level
$\alpha$ defined by the `make_coefficients` function, plus a linear combination
of the environmental variables. For efficiency, this linear combination is 
calculated with matrix multiplication. The weights associated with this linear 
combination (`alpha_env`) are drawn from a Gaussian distribution with zero mean 
and unit variance.

For the abundance-based simulations, each species' conditional abundance was
simulated with a Poisson distribution instead of a Bernoulli distribution. The
rate parameter of this distribution was given by 
$\mathrm{log}(1 + exp(\alpha_i + \sum_j\beta_{ij}y_j))$. (The more conventional 
exponential inverse link was not used because preliminary experiments showed that
it led to very large abundances that prevented the Gibbs sampler from mixing 
effectively.)

Each simulated landscape was saved after the Gibbs sampler had updated the 
value of each species `n_gibbs` times; see the end of this document for the
value used.

@@

  # Landscape starts as if betas were all zero. Each species' occurrence probability
  # or abundance depends on its alpha value and on the environment (assuming alpha_env 
  # is not set to zero).
  x = matrix(
    f(rep(1, n_sites) %*% t(alpha) + env %*% alpha_env), 
    nrow = n_sites, 
    ncol = n_spp
  )

@@

Loop to simulate and save 50 landscapes for each combination of simulation type
and landscape size.
